---
title: "STAA 566 Dynamic Graphs Assignment"
author: "Ruth Hilton"
output: html_document
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
---

For this assignment, I have chosen to investigate results data from the International Mathematics Olympiad from 1984-2017. This eaier-to-use data set was downloaded from Kaggle.com, but the full results can also be accessed directly imo-official.com. This data set includes the names, scores, ranks and awards of each individual who participated in the IMO from 1984 until 2017. 

```{r message = FALSE}
#load libraries
library(dplyr)
library(ggplot2)
library(ggthemes)
library(plotly)
library(htmlwidgets)
library(viridis)
```

First I will create a plot of the highest scorer each year against the competition year for both the USA and Germany, without any dynamic elements.

```{r fig.height = 3, fig.width = 5}
imo.results <- read.csv('imo_results.csv') #read data file
plot.USA <- imo.results %>% 
  filter(country == "USA") %>% #filter data to USA participants
  group_by(year) %>% 
  slice(which.max(total)) %>% #choose the highest scorer from the USA each year
  ggplot(aes(x = year, 
             y = total,
             color = rank,
             text1 = firstname,
             text2 = lastname)) +
  geom_point(shape = 16) +
  theme_clean() +
  scale_color_viridis() +
  xlab(NULL) + 
  ylab("Total Point Score (0-42)") + 
  ggtitle("Highest Scoring Americans at International Mathematics Olympiad", 
          subtitle = "Competition Years 1984-2017") +
  scale_x_continuous(breaks = seq(1984, 2017,  by = 3))

plot.GER <- imo.results %>% 
  filter(country == "GER") %>% #filter data to GER participants
  group_by(year) %>% 
  slice(which.max(total)) %>% #choose the highest scorer from GER each year
  ggplot(aes(x = year, 
             y = total,
             color = rank,
             text1 = firstname,
             text2 = lastname)) +
  geom_point(shape = 16) +
  theme_clean() +
  scale_color_viridis() +
  xlab(NULL) + 
  ylab("Total Point Score (0-42)") + 
  ggtitle("Highest Scoring Germans at International Mathematics Olympiad", 
          subtitle = "Competition Years 1984-2017") +
  scale_x_continuous(breaks = seq(1984, 2017,  by = 3))
  
plot.USA
plot.GER
```

Now I will use `plotly` to add interactive elements to this graphic and display the two plots side by side.

```{r}
dplot.USA <- ggplotly(plot.USA,
                  tooltip = c("text1", "text2", "y", "colour"))
dplot.GER <- ggplotly(plot.GER,
                  tooltip = c("text1", "text2", "y", "colour"))
subplot(dplot.GER, dplot.USA, shareY = TRUE) %>% layout(title ="High Scorers at IMO from Germany and the USA")
```

*Data Source: https://www.kaggle.com/luckyt/imo-scores*